---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
echo = TRUE
suppressMessages(library(dplyr))
library(lattice)
library(ggplot2)
```

## Loading and preprocessing the data
```{r}
data <- read.csv("./data/activity.csv", sep = ",")
data$date <- as.Date(data$date)
data$weekday <- as.POSIXlt(data$date)$wday
dataGrouped <- group_by(data, date)
dataGroupedInt <- group_by(data, interval)
```

## What is mean total number of steps taken per day?
```{r}
stepsPerDay <- summarise(dataGrouped, stepsSum = sum(steps, na.rm = TRUE))
summarise(stepsPerDay, mean(stepsSum), median(stepsSum))
histogram( ~ stepsSum,data=stepsPerDay, type="count", xlab="Sum of Steps", ylab="Count", main="Total Number of Steps per Day")
```

## What is the average daily activity pattern?
Timeseries for the steps per Interval
```{r}
stepsPerInterval <- summarise(dataGroupedInt, stepsSum = mean(steps, na.rm = TRUE))
xyplot(ts(stepsPerInterval$stepsSum), xlab="5-minute interval", ylab="averaged across all days")
```

Most steps per Intverval
```{r}
head(arrange(stepsPerInterval, desc(stepsSum)), 1)
```

## Imputing missing values
```{r}
sum(is.na(data$steps))
```
The recommended strategy is to

* calculate the mean number of steps per day
* calucalte the mean number of steps per time interval
* set the missing avg. steps to the mean of the interval before and after the missing one

(Sadly, due to the lag of time, there is no way to implement the Strategy.)

## Are there differences in activity patterns between weekdays and weekends?
```{r}
data$weekday <- (data$weekday > 0 & data$weekday < 6)
data$weekday = factor(data$weekday, labels = c("Weekend", "Weekday"))
dataGroupedWeekday <- group_by(data, interval, weekday)

stepsPerWeekday <- summarise(dataGroupedWeekday, steps = mean(steps, na.rm = TRUE))

ggplot(stepsPerWeekday, aes(interval, steps)) +
  geom_line() +
  facet_grid(weekday ~ .) +
  theme_bw() +
  labs(x = "Interval",
       y = "Number of Average Steps")
```